### 2021-01-31 research project/sections/Analysis_of_results.tex
rp work
rp

The arithmetic mean of the 100 trials was caluclated and the data points 
in the table shown below were found (rounded to 3 s.f.):
\begin{table}[h]
    \noindent% <-- important
    \setlength\tabcolsep{3pt} % default: 6pt
    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
    \toprule
    Decimal place & Average time for approximation using Viete's method, in milliseconds (ms) & Average time for approximation using Madhava's method method, in milliseconds (ms) \\ 
    \midrule
    5  & 0.760  & 0.592   \\
    10 & 1.29  & 1.31    \\
    15 & 2.05  & 1.9     \\
    20 & 2.64  & 2.68    \\
    25 & 3.09  & 3.21    \\
    30 & 3.82  & 3.9     \\
    35 & 4.44  & 4.55    \\
    40 & 5.08  & 5.26    \\
    45 & 5.71  & 5.84    \\
    50 & 6.37  & 6.58    \\
    55 & 7.02  & 7.27    \\
    60 & 7.72  & 8.00   
    \end{tabularx}
\end{table}

@@

To better show the trend in the data collected, the data 
has been presented in a chart below. The blue markers and line of best-fit 
respresent the results received from Viète's method while the red represent

@@

>>>>

The arithmetic mean of a decimal place was caluclated for 100 trials and the data points 
in the table shown below were found (rounded to 3 s.f.):
\begin{table}[h]
    \noindent% <-- important
    \setlength\tabcolsep{3pt} % default: 6pt
    \begin{tabularx}{\textwidth}{@{} l *{6}{C} @{}}
    \toprule
    Decimal place & Average time for approximation using Viete's method, in milliseconds (ms) & Average time for approximation using Madhava's method method, in milliseconds (ms) \\ 
    \midrule
    5  & 0.760  & 0.592   \\
    10 & 1.29  & 1.31    \\
    15 & 2.05  & 1.9     \\
    20 & 2.64  & 2.68    \\
    25 & 3.09  & 3.21    \\
    30 & 3.82  & 3.9     \\
    35 & 4.44  & 4.55    \\
    40 & 5.08  & 5.26    \\
    45 & 5.71  & 5.84    \\
    50 & 6.37  & 6.58    \\
    55 & 7.02  & 7.27    \\
    60 & 7.72  & 8.00   
    \end{tabularx}
\end{table}

@@

To better show the trend in the data collected, the data 
has been presented in a chart below. The blue markers and line of best-fit 
respresent the results received from Viète's method while the red represent those 
from Madhava's method.

@@

The results gathered from this experiment are descisive.

### 2021-01-31 research project/sections/Background.tex
rp

In this paper, we focus on two methods for approximating the value of $\pi$. 
The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama and 
rediscovered by Swiss mathematician Leibniz. Two methods with different approaches have been chosen for 
comparison, the process for the original discovery of these methods
will be explained. The two methods in question are one based on 
the infinite series definition of an inverse trigonometric function 
and one where the value is derived using geometry, by mathematician 
Viète. These two mathematical methods were chosen as they are both 
represent a first occurence in mathematics: Madhava was the first to 
find the series notation of the inverse tangent function and Viète was 
the first mathematician to use and infinite product in calculation.

\subsubsection{Madhava-Leibniz method}

Madhava is the first mathematician known to have found the series notation for 
the inverse tangent function, in the Kerala school of mathematics of Medieval India 
\cite{amermathmonthly}. One of

\subsubsection{Viète's method}

François Viète approached the value of $\pi$ from a 
geometric standpoint, and found an infinite product. 
He was able to calculate $\pi$ to a place of 9 decimal points, in 
the year 1593 \cite{Kreminski}, using his method. His method is reminiscent 
of Archimedes' method, where the length of a side is calculated \cite{archimedespi}, 
but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant 
radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon 
tends toward the area of a circle. The geometric origin of this formula 
can be found using simple right-angle trigonometry, by first finding 
the lengths $OH$ and subsequently $BD$ in \ref{fig:image1}.
% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
With the radius of the circle with center $R = OB$,

@@

So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and 
followingly, it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The 
value of $R = 1$ in this case, and since the area of a circle is defined 
by $A = \pi R^2$, therefore: \\
$\pi = \frac{A_{n}}{P}$ \\
By definition, we can say that as the value $k$ approaches infinity, the area of the $kn$-gon 
aproaches that of a circle, and therefore, the value of $\pi$.

>>>>

In this paper, we focus on two methods for approximating the value of $\pi$. 
The first by French mathematician François Viète, and another discovered by Madhava of Sangamagrama and 
rediscovered more than a hundred years later by James Gregory. Two methods with different 
approaches have been chosen for comparison and the process for original discovery of these methods
will be explained. The two in question are one based on 
the infinite series definition of an inverse trigonometric function discovered by Madhava
and one where the value is derived using geometry, by mathematician 
Viète. These two mathematical methods were chosen as they are both 
represent a first occurrence in mathematics: Madhava was the first to 
find the infinite series expansion of the inverse tangent function and Viète was 
the first mathematician to use an infinite product in calculation.

\subsubsection{An analytical method: Madhava's method}

Madhava is the first mathematician known to have found the series notation for 
the inverse tangent function, in the Kerala school of mathematics of Medieval India 
\cite{amermathmonthly}. The research made at the Kerala school was documented by 
many mathematicians of their time, namely the astronomer-mathematician Jyesthadeva 
in his treatise \textit{Yukti-Bhasa} written in the Malayalam language \cite{oconnor_robertson_2000}. 
Madhava's infinite series expansion of the inverse tangent function can also be 
found in this treatise. R. C. Gupta, in his translation, describes this expansion as one 
where the arc $\theta$ is equal to the sum of a first term, "the product of the given sine 
and radius of the desired arc divided by the cosine of the arc" \footnotemark, followed by terms that "are 
obtained by a process of iteration": . \cite{rc_gupta_mgseries}

\footnotetext{Note: the old Indian meaning for the sine of $\theta$ is $r \sin{\theta}$, where $r$ is the radius. The same applies for the cosine.}

This text explains what can be written in modern mathematical terms as such:

$r \theta = \frac{r (r \sin{\theta})}{1 (r \cos{\theta})} - \frac{r (r \sin{\theta})^3}{3 (r \cos{\theta})^3} + \frac{r (r \sin{\theta})^5}{5 (r \cos{\theta})^5} - \frac{r (r \sin{\theta})^7}{7 (r \cos{\theta})^7} + \dots$

And for a circle of radius $r = 1$, we can cancel all $r$ terms:

$\theta = \frac{\sin{\theta}}{\cos{\theta}} - \frac{\sin^3{\theta}}{3 \cos^3{\theta}} + \frac{\sin^5{\theta}}{5 \cos^5{\theta}} - \frac{\sin^7{\theta}}{7 \cos^7{\theta}} + \dots$

And because $\frac{\sin{\theta}}{\cos{\theta}} = \tan{\theta}$, the aforementioned expression 
can be simplified to:

$\theta = \tan{\theta} - \frac{\tan^3{\theta}}{3} + \frac{\tan^5{\theta}}{5} - \frac{\tan^7{\theta}}{7} + \dots$

So then, if we let $\tan{\theta} = \alpha$, we can find the infinite series expansion 
of the arctangent function.

$\arctan{\alpha} = \alpha - \frac{\alpha^3}{3} + \frac{\alpha^5}{5} - \frac{\alpha^7}{7} + \dots = \theta$

Moreover, it is said in a further commentary describing the findings of Madhava that a  
method for the calculation of the circumference $c$ of a circle of diameter $d$ exists. 
The passage in question from the commentary \textit{Tantrasamgraha-vyakhya} of 
anonymous authorship states that, by following the same argument as stated before, one 
can find the circumference of a circle through a similar infinite sum. The first 
term of this sum would be "the square root of the square of the diameter multiplied by 
twelve", followed by the first term "divided by three in each successive case", and 
when these "are divided in order by the odd numbers, beginning with 1", and after the 
even terms are subtracted from sum of the odd", one is left with the circumference of the 
circle (translation from C. K. Raju). \cite{raju_2007}

This can be expressed in mathematical terms as such:

\subsubsection{A geometric method: Viète's method}

François Viète approached the value of $\pi$ from a 
geometric standpoint, and found an infinite product. 
He was able to calculate $\pi$ to a place of 9 decimal points, in 
the year 1593 \cite{Kreminski}, using his method. His method is reminiscent 
of Archimedes' method, where the length of a side is calculated \cite{archimedespi}, 
but differs in that it consists of finding the area of a polygon of $n$ sides in a circle of constant 
radius, rather than the circumference. As the value of $n$ is increased, the area of the $n$-gon 
tends toward the area of a circle. The geometric origin of this formula 
can be found using simple right-angle trigonometry, by first finding 
the lengths $OH$ and subsequently $BD$ in (Figure 1 ) \ref{fig:image1}.
% $$\pi = \frac{2}{\prod_{k = 0}^{\infty} U_{k}}$$    symbolic
With the radius of the circle with center $R = OB$,

@@

So $P = \frac{A_{n}}{A}$, since the values $A_{kn}$ cancel, and 
so it is true that $P = \frac{A_{n}}{A} \Leftrightarrow A = \frac{A_{n}}{P}$. The 
value of $R = 1$ in this case, and since the area of a circle is defined 
by $A = \pi R^2$, therefore: \\
$\pi = \frac{A_{n}}{P}$ \\
By definition, we can say that as the value $k$ approaches infinity, the area of the $kn$-gon 
approaches that of a circle, and therefore, the value of $\pi$.
